Separating Use and Reuse to Improve Both

### 2017-05-23 FACS/section4.tex
Section 4

\begin{lstlisting}
  CPoint:Use pointSum,pointMul, {//not our suggested solution
    int x   int y
    method int x() x       
    method int y() y
    class method This of(int x, int y) new Point(x,y)
    constructor Point(int x, int y){ this.x=x   this.y=y }
    }
\end{lstlisting}

@@

\begin{lstlisting}
  PointAlgebra:Sum pointSum,pointMul
\end{lstlisting}

  As we can see, we can declare the methods independently and compose the result
  as we wish. In this first attempt, we repeated the abstract methods \Q@x@,\Q@y@ and \Q@of@.
  In addition to \Q@sum@ and \Q@mul@ we may want many operations over points; can we improve our reuse
  and not repeat such abstract definitions? of course!

@@

Now our code is fully modularized, and each trait handle exactly one method.

\paragraph{Adding state:}${}_{}$\\
What happens if we want to add fields instead of just operations?

@@

This first attempt does not work: the abstract color method
is not a getter for any of the parameters of 
\Q@ class method This of(int x,int y)@
A solution is to provide a richer factory:

@@

where we assume to support overloading on different parameter number.
This is a good solution, indeed we think is better that any alternatives in literature,
however the method \Q@CPoint.sum@ resets the color to red: we call the \Q@of(int,int)@ method, that now delegate to \Q@of(int,int,Color)@ by passing red as the default field value.
What should be the behaviour in this case?
If our abstract state supports withers, instead of writing \Q@This.of(...)@ we can use
\Q@this.withX(newX).withY(newY)@ in order to preserve the color from \Q@this@.
This solution is still not sadisfactory: this design ignore the color from \Q@that@.

If the point designer could predict this kind of extension, then we could use the following better design:  
\begin{lstlisting}
  p: { method int x() method int y() //getters
    method This withX(int that) method This withY(int that)//withers
    class method This of(int x,int y)
    method This merge(This that) //new method merge!
    }
  pointSum:Use p, { method This sum(This that)
      this.merge(that).withX(this.x()+that.x()).withY(this.y()+that.y())
    }
  colored:{method Color color()
    method This withColor(Color that)
    method This merge(This that) //how to merge color handled here
      this.withColor(this.color().mix(that.color())
    }
  CPoint:/*as before*/
\end{lstlisting}  
  Now we can merge colors, or any other kind of state we may want to add
  following this pattern.

  In order to compose, let say \Q@colored@ with \Q@flavored@ we would
  need to compose the merge operation inside of both of them.
  The simple model we are presenting could accomodate this with an
  extension allowing code literals inside of a \use\ expression to use some form of super call to compose conflicting implementations. This is similar to the \emph{override} operation present in the original trait model~\cite{ducasse2006traits}.

>>>>

\begin{lstlisting}
  CPoint:Use pointSum,pointMul, {//not our suggested solution
    int x   int y
    method int x() x       
    method int y() y
    class method This of(int x, int y) new Point(x,y)
    constructor Point(int x, int y){ this.x=x   this.y=y }
    }
\end{lstlisting}
\bruno{We talk about withers later on. So I think we should consider
  having withers in this code, so that readers can understand what 
withers are!}

@@

\begin{lstlisting}
  PointAlgebra:Use pointSum,pointMul
\end{lstlisting}

  As we can see, we can declare the methods independently and compose the result
  as we wish.

  \paragraph{Improved solution} In this first attempt, we repeated the abstract methods \Q@x@,\Q@y@ and \Q@of@.
  Moreover, in addition to \Q@sum@ and \Q@mul@ we may want many
  operations over points. It is possible to improve reuse
  and not repeat such abstract definitions by abstracting the common
  abstract definitions into a trait \Q@p@:

@@

Now the code is fully modularized, and each trait handles exactly one method.

\subsection{State Extensibility}
Programmers may want to extend points with more state. For example 
they may want to add colors to the points. A first attempt at doing
this would be:

@@

This first attempt does not work: the abstract color method
is not a getter for any of the parameters of 
\Q@ class method This of(int x,int y)@. 
A solution is to provide a richer factory:

@@

\noindent 
where we assume support for overloading on different number of parameters.
This is a reasonable solution, however the method \Q@CPoint.sum@ resets
the color to red: we call the \Q@of(int,int)@ method, that now
delegates to \Q@of(int,int,Color)@ by passing red as the default field
value.  What should be the behaviour in this case?  If our abstract
state supports withers, instead of writing \Q@This.of(...)@ we can use
\Q@this.withX(newX).withY(newY)@ in order to preserve the color from
\Q@this@.  This solution is still not satisfactory: this design ignores
the color from \Q@that@.

\paragraph{A better design}
If the point designer is designing for reuse and extensibility, then 
a better design would be the following:  
\begin{lstlisting}
  p: { method int x() method int y() //getters
    method This withX(int that) method This withY(int that)//withers
    class method This of(int x,int y)
    method This merge(This that) //new method merge!
    }
  pointSum:Use p, { method This sum(This that)
      this.merge(that).withX(this.x()+that.x()).withY(this.y()+that.y())
    }
  colored:{method Color color()
    method This withColor(Color that)
    method This merge(This that) //how to merge color handled here
      this.withColor(this.color().mix(that.color())
    }
  CPoint:/*as before*/
\end{lstlisting}  
  This design allows merging colors, or any other kind of state we may want to add
  following this pattern.\bruno{worried that withers are not explained
  enough.}

\paragraph{Independent Extensibility}
  Ofcourse, quite frequentely there can be multiple independent
  extensions~\cite{Zenger-Odersky2005} that need to be composed. Lets suppose that 
  we could have a notion of flavoured points as well.   
  In order to compose, let say \Q@colored@ with \Q@flavored@ we would
  need to compose the merge operation inside of both of them.
  The simple model we are presenting could accomodate this with an
  extension allowing code literals inside of a \use\ expression to use some form of super call to compose conflicting implementations. This is similar to the \emph{override} operation present in the original trait model~\cite{ducasse2006traits}.
